The Astropy Project

### 2017-11-08 main.tex
Update on Overleaf.

\section*{\textit{Notes and guidelines (to be removed)}}

@@

Central to the success of \astropy is an open environment where
anybody can contribute to the code by opening a so-called ``pull
request'' with a new code contribution on a publically available
website\footnote{\url{https://github.com/astropy/astropy}}; this
request is then reviewed by at least one of the maintainers responsible for the specific
sub-package and possibly merged into the main code base. Most bug
fixes and small features are contributed in this way. However, this
model leads to ``organic'' growth where different features are
implemented by different people with different programming styles and
interfaces. Thus, \astropy has a mechanism to more formally propose
changes to the code management (e.g.~the release cycle) or to plan
out major new features. This mechanism is called ``Astropy Proposal
for Enhancement'' (APE). In an APE one or more authors lay out a
vision for a new feature, including a rationale why this feature is
needed, how it should be implemented, and what the interface will
be. These APEs are discussed and refined before much work is invested
into a detailed implementation. APEs are again edited and discussed on
a platform\footnote{\url{https://github.com/astropy/astropy-APEs}}
open to the public. In previous APEs a community consensus emerged and
APEs are accepted and become the basis for future work at this
point. In cases where consensus cannot be reached, the
\astropy coordination committee can decide to close the discussion and
make an executive decision based on the community input on the APE in
question. APEs are modeled after the ``Python Enhancement Proposals''
(PEP) and guide the development of the \python programming language.

@@

\subsection{Release cycle and Long Term Support}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
\inlinecomment{BMS: TODO}{reference APE2 for versioning and APE10 for dropping python2 support}
\inlinecomment{TR}{added a reference to APE10, will add APE2 once we upload that one to Zenodo}

\inlinecomment{BMS}{shall we maybe refer to the release calendar, too?}
\inlinecomment{hamogu}{I don't think that's necessary. These details are in the docs, but not too important for a general paper.}
\inlinecomment{BMS}{should we mention that v3.0 is the exception to the
  version numbering scheme we use by not being 2.1?}
\inlinecomment{hamogu}{No. This is a genral paper, those are details that will be unimportant very soon. }
%
\par The general scheme for the releases and version numbering that the core
package has adopted is the following. \astropypkg has new feature releases every
6 months, and between feature releases additional bugfix releases that
contain only bugfixes but no new features.
% The timing of
%
\par Long-term support (LTS) releases continue to receive bugfixes for 2
years with no changes to the API\@. They are ideal for pipelines and other
applications where API stability is essential. The latest LTS release is
also the last one that supports \python 2, and will receive bug fixes until the
end of 2019 \citep{ape10}.
%
\par The version numbering of \astropypkg reflects on the release scheme
described above. The core package uses the form x.y.z, where x is advanced
on LTS, y is advanced for a non-LTS feature release, and z is advanced on
bugfix releases.

@@

\subsubsection{Overview}
The \package{astropy.coordinates} subpackage is designed to support representing
and transforming celestial coordinates and, new in version 2.0, velocities.
The framework heavily relies on the \package{astropy.units} subpackage, and most
inputs to objects in this subpackage are expected to be \texttt{Quantity}
objects.
Some of the machinery also relies on the Essential Routines of Fundamental
Astronomy (ERFA) \texttt{C} library for some of the critical underlying
transformation machinery \citep{erfa}, which is based on the Standards Of
Fundamental Astronomy (SOFA) effort \citep{sofa}.

@@

\inlinecomment{Should we mention that Cutout2D supports WCS to pixel conversions?}

\inlinecomment{Yes, done.}

@@

\inlinecomment{Tom A.}

@@

\inlinecomment{SCO}{Pyfits deprecation in favor of io.fits ?}
\inlinecomment{SCO}{Table and FITS tables}

\subsubsection{Unified file read/write interface}

TODO: Check when it was added (0.3?). Overview of the supported formats (ASCII,
FITS, votable, HTML, JSViewer/Datatables).

@@

\inlinecomment{SCO}{One subsection per format (if enough new features) ?}
\inlinecomment{BMS}{Do we really need to mention the old version numbers in the list below when a feature was added?}

@@

    \item[ASCII: fast reader/writer (v1.0)]

@@

    \item[ASCII: HTML tables (v0.4)]

@@

	\item [Enhanced CSV format (v1.0)]

@@

	\item [FITS: command-line scripts (\texttt{fitsheader}, \texttt{fitsinfo}, v0.4)]

@@

	\item [FITS: lazy loading (v1.3)]

@@

	\item [Misc: YAML serialization (v1.3)]

@@

\inlinecomment{BMS}{Maybe move the command line tools into a separate subsection to highlight them (even though we mostly only have fits related scripts)? They are supposed to be used more frequently by individual users.}
\inlinecomment{hamogu}{MY feeling is that this section is too detailed already. If I had written it, I would not have mentioned the scripts at all.}

@@

\cite{Lupton2004} describe an ``optimal'' algorithm for producing red-green-blue (RGB) composite images from three separate high-dynamic range arrays. The \package{visualization} package provides a convenience function to create such a color image.  It also includes an associated set of classes to provide alternate scalings.
This functionality was contributed by developers from the Large Synoptic Survey Telescope (LSST) and serves as an example of contribution to Astropy from a more traditional engineering organization \citep{lsst_astropy}.

@@

Development has continued after the release of v2.0 of \package{astropy}.    The most significant development for the next release will be removing python 2 support \cite{python 3 APE}.   While the v2.0 release is a ‘long term stable’ release )meaning that it will be supported through 2019 in both python 2 and python 3), the v3.0 \package{astropy} release will only support python 3.  This will allow the use of the new functionality in python 3, simplify the code base, and reducing the testing overhead.   The v3.0 release is currently scheduled for December 2017.

>>>>

l\section*{\textit{Notes and guidelines (to be removed)}}

@@

\inlinecomment{Tom R}{at the moment this section doesn't take into account that APEs can be unrelated to code (e.g. APE 8 for the code of conduct)}

Central to the success of \astropy is an open environment where
anybody can contribute to the code by opening a so-called ``pull
request'' with a new code contribution on a publically available
website\footnote{\url{https://github.com/astropy/astropy}}; this
request is then reviewed by at least one of the maintainers responsible for the specific
sub-package and possibly merged into the main code base. Most bug
fixes and small features are contributed in this way. However, this
model leads to ``organic'' growth where different features are
implemented by different people with different programming styles and
interfaces. Thus, \astropy has a mechanism to more formally propose
changes to the code management (e.g.~the release cycle) or to plan
out major new features. This mechanism is called ``Astropy Proposal
for Enhancement'' (APE). In an APE one or more authors lay out a
vision for a new feature, including a rationale why this feature is
needed, how it should be implemented, and what the interface will
be \citep{ape1}. These APEs are discussed and refined before much work is invested
into a detailed implementation. APEs are again edited and discussed on
a platform\footnote{\url{https://github.com/astropy/astropy-APEs}}
open to the public. In previous APEs a community consensus emerged and
APEs are accepted and become the basis for future work at this
point. In cases where consensus cannot be reached, the
\astropy coordination committee can decide to close the discussion and
make an executive decision based on the community input on the APE in
question. APEs are modeled after the ``Python Enhancement Proposals''
(PEP) and guide the development of the \python programming language.

@@

\subsection{Release cycle and Long Term Support}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
\inlinecomment{BMS}{shall we maybe refer to the release calendar, too?}
\inlinecomment{hamogu}{I don't think that's necessary. These details are in the docs, but not too important for a general paper.}
\inlinecomment{BMS}{should we mention that v3.0 is the exception to the
  version numbering scheme we use by not being 2.1?}
\inlinecomment{hamogu}{No. This is a genral paper, those are details that will be unimportant very soon. }
%
\par The general scheme for the releases and version numbering that the core
package has adopted is the following. \astropypkg has new feature releases every
6 months, and between feature releases additional bugfix releases that
contain only bugfixes but no new features \citep{ape2}.
% The timing of
%
\par Long-term support (LTS) releases continue to receive bugfixes for 2
years with no changes to the API\@. They are ideal for pipelines and other
applications where API stability is essential. The latest LTS release is
also the last one that supports \python 2, and will receive bug fixes until the
end of 2019 \citep{ape10}.
%
\par The version numbering of \astropypkg reflects on the release scheme
described above. The core package uses the form x.y.z, where x is advanced
on LTS, y is advanced for a non-LTS feature release, and z is advanced on
bugfix releases.

@@

\subsubsection{Overview}
The \package{astropy.coordinates} subpackage is designed to support representing
and transforming celestial coordinates and, new in version 2.0, velocities.
The framework heavily relies on the \package{astropy.units} subpackage, and most
inputs to objects in this subpackage are expected to be \texttt{Quantity}
objects.
Some of the machinery also relies on the Essential Routines of Fundamental
Astronomy (ERFA) \texttt{C} library for some of the critical underlying
transformation machinery \citep{erfa}, which is based on the Standards Of
Fundamental Astronomy (SOFA) effort \citep{sofa}.
% timj: The preferred SOFA reference is usually https://ui.adsabs.harvard.edu/#abs/2011SchpJ...611404H/abstract rather than a URL to the SOFA web site.

@@

@@

@@

%\inlinecomment{SCO}{Pyfits deprecation in favor of io.fits ?}
%\inlinecomment{SCO}{Table and FITS tables}

%\subsubsection{Unified file read/write interface}

%\inlinecomment{Check when it was added (0.3?). Overview of the supported formats (ASCII,FITS, votable, HTML, JSViewer/Datatables).}

@@

@@

    \item[ASCII: fast reader/writer]

@@

    \item[ASCII: HTML tables]

@@

	\item [Enhanced CSV format ]

@@

	\item [FITS: command-line scripts (\texttt{fitsheader}, \texttt{fitsinfo})]

@@

	\item [FITS: lazy loading ]

@@

	\item [Misc: YAML serialization ]

@@

%\inlinecomment{BMS}{Maybe move the command line tools into a separate subsection to highlight them (even though we mostly only have fits related scripts)? They are supposed to be used more frequently by individual users.}
%\inlinecomment{hamogu}{MY feeling is that this section is too detailed already. If I had written it, I would not have mentioned the scripts at all.}

@@

\cite{Lupton2004} describe an ``optimal'' algorithm for producing red-green-blue (RGB) composite images from three separate high-dynamic range arrays. The \package{visualization} package provides a convenience function to create such a color image.  It also includes an associated set of classes to provide alternate scalings.
This functionality was contributed by developers from the Large Synoptic Survey Telescope (LSST) and serves as an example of contribution to Astropy from a more traditional engineering organization \citep{lsst_astropy}.
% Please cite the final SPIE paper here rather than the presentation slides at that conference.
% https://ui.adsabs.harvard.edu/#abs/2016SPIE.9913E..0GJ/abstract (I'm not sure how to edit the bib file on overleaf)

@@

Development has continued after the release of v2.0 of \package{astropy}.    The most significant development for the next release will be removing python 2 support \cite{ape10}.   While the v2.0 release is a ‘long term stable’ release )meaning that it will be supported through 2019 in both python 2 and python 3), the v3.0 \package{astropy} release will only support python 3.  This will allow the use of the new functionality in python 3, simplify the code base, and reducing the testing overhead.   The v3.0 release is currently scheduled for December 2017.
